---
title: "Proportions Less Than 1/n"
author: "Eric C. Anderson"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: true
    toc: true
  bookdown::html_document2:
    base_format: rmarkdown::html_vignette
    fig_caption: true
    toc: true
  html_document:
    df_print: paged
    toc: true
  word_document: default
bibliography: ../../vignettes/citation.bib
csl: ../../vignettes/ecology.csl
---


\newcommand{\bpi}{\boldsymbol{\pi}}


```{r setup, include=FALSE}
library(tidyverse)
```


## Introduction 

One of the main outputs from 'rubias' is a Markov chain Monte Carlo (MCMC) sample of
the mixing proportion.  This is meant to be a sample from the posterior distribution
of the vector of mixing proportions from which the sample was drawn.  This sample is
returned from the `infer_mixture()` function in the `mix_prop_traces` component,
of the return object, and the posterior mean of the mixing proportions is returned
in the `mixing_proportions` component.  

During each step of the MCMC the current value of the mixing proportions are drawn
as a sample from a Dirichlet distribution.  This is a vector $\bpi = (\pi_1,\ldots,\pi_K)$
of values with $0< \pi_k < 1$ for all $k$ and $\sum_{k=1}^K \pi_k = 1$.  The individual
$\pi_k$ values are constrained to not be equal to 0 or 1.  Thus, even if there are no fish
from a certain collection in the mixture, there might still be a small, but non-zero,
estimate of the proportion of fish from that collection.  Depending on they are
interpreted, these small, but non-zero estimates can lead to different thoughts
about how fisheries should be managed.  For example, if a very small proportion is
interpreted as evidence that there is a non-zero chance that a rare or threatened
stock is being impacted by a fishery, then some might wish to curtail the fishery
on that basis, even if none of the fish in the sample are actually from that
threatened stock. If it actually is the case that none of the fish in the sample
are from the threatened stock, then fishery restrictions put in place simply because
of a non-zero estimate of the mixing proportion would incur a loss (lost income,
reduced catch, etc.); however since no fish from the threatened stock were in the
sample, it is far from clear that the fishery restrictions would actually be
benefitting the threatened stock. 


Similar dynamics can affect the output of the newly implemented approach in 'rubias'
to quantify the uncertainty on the total catch from the genetic data. (See the
[vignette](https://eriqande.github.io/rubias/articles/uncertainty-on-total-catch.html) about
this.)  In that work, the actual allocations of fish in the sample to different collections
are used to simulate from the posterior predictive distribution of a much larger sample.
In this case, especially if the catch is being expanded to the numbers of fish that
are in the escapement, for example, the occasional allocation of a fish to a rare
stock can lead to a large estimate (with considerable variance) of the number of
fish returning to that stock.  In this case it would be natural to ask whether
there was high confidence in the allocation of that fish to the rare stock. 

There are reports of small, non-zero estimates of mixing proportions that are
considerably less than $1/N$, where $N$ is the number of fish sampled from the
mixture.  That would indicate that there is not even a single fish that can be
said to confidently belong to the population.  There might also be estimated proportions
that are greater than $1/N$, but which are based on a handful of fish, none of
possess compelling genetic data that they are actually from the stock in question.

I have been thinking of ways to
approach this problem so that such small, but non-zero, proportions could be
given less weight, or at least could be identified. This might be a
good candidate for a "spike-and-slab" type of prior, that is hoped to make the
Dirichlet prior parameter on truly empty components a lot smaller than it is for
components that are not empty.  Or, perhaps the problem can be better approached
in a decision-theoretic framework applied to the currently calculated posterior.

Regardless of how we decide to approacy the problem, I think it would be
good to have some tools to really let us explore the details of this problem. I
have a lot of questions: is the small non-zero weight coming entirely from the
prior, or are there some fish that are occasionally being assigned to that
component? Is it just one fish, or is a little bit of weight from many fish?
Do those fish suffer from lots of missing data so that they are getting smeared
across a lot of different populations?  These are all things that would be good
to know and to be able to diagnose, whether or not we have a formal
solution to this problem.  Furthermore, whatever solution we do come up with,
any tools to explore the problem now will likely be useful for assessing how well
the solution works.  


## Assessment Tools

First and foremost, when assessing these sorts of issues, it would be handy to have
a record of the actual full-conditional probabilities for each fish at each
sweep of the MCMC algorithm.  These full-conditional probabilities are effectively
the posterior probabilities of assignment for each fish given the current values
of the mixing proportions ($\bpi$).  We never included code to output these, because
the size of the output might be rather larger.  However, in the case of the default
conditional mixture model, (`infer_mixture()` with `method="MCMC"`) the actual
full_conditional probabilities can be reconstructed using only the sampled $\bpi$
values---which are already returned in `$mix_prop_traces` and the matrix of scaled
likelihoods.  The scaled likelihoods have not been traditionally returned by
`infer_mixture()`, but I have updated the function so that it always returns
the scaled likelihoods in a matrix.  This can be handy in a number of ways for
the problem at hand.  


## A Cool plot

Let's get some example results:
```{r}
library(tidyverse)

mixf <- infer_mixture(chinook, chinook_mix, 5)
```


