---
title: "# rubias --- a package for bias correction in hierarchical GSI"
date: "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: rubias_bib.bibtex
output: 
  github_document:
    toc: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "readme-figs/"
)
library(tidyverse)
```




This is an R package designed to diagnose and correct a bias recently 
documented in genetic stock identification. The bias occurs when mixture 
proportion estimates are desired for groups of populations (reporting units)
and the number of populations within each reporting unit are uneven.

In order to run C++ implementations of MCMC, rubias requires the package
Rcpp, which in turn requires an Rtools installation. After cloning into the
repository with the above dependencies installed, build & reload the package 
to view further documentation.

The script "/R-main/coalescent_sim" was used to generate coalescent simulations
for bias correction validation. This is unnecessary for testing the applicability
of our methods to any particular dataset, which should be done using the 
Hasselman_simulation_pipeline and/or bias_comparison functions.
coalescent_sim creates simulated populations using the ms coalescent simulation 
program, available from the Hudson lab at UChicago, and the GSImulator
and ms2geno packages, available at https://github.com/eriqande, and so requires
further dependencies than the rest of the package.


# Input Data

The functions for conducting genetic mixture analysis and for doing simulation
assessment to predict the accuracy of a set of genetic markers for genetic stock 
identification require that genetic data be input as a data frame in a
specific format:

- one row per individual
- each locus is represented by two adjacent columns, one for each allele (this package is only
  configured for diploids, at the moment). Allelic types can be expressed as any number
  or character
- missing data at a locus is expressed with NA values for each gene copy at the locus
- the name of the locus is taken to be the column name of the first column of each pair
of locus columns.  The header on the second column is ignored.
- the data frame must have four columns of meta data for each individual:
    * `sample_type`:  a column telling whether the sample is a `reference` sample or a `mixture` sample.
    * `repunit`: the reporting unit that an individual belongs to. This is required if sample_type is `reference`.
       And if  sample_type  is `mixture` then repunit must be `NA`.  
    * `collection`: for reference samples, the name of the population that the individual is from. For mixture
    samples, this is the name of the particular sample (i.e. stratum or port that is to be treated together in 
    space and time.)
    * `indiv` a character vector with the ID of the fish.  These must be unique.
- Note that both the `repunit` and the `collection` columns can either be character vectors or, if a particular sort order is desired, they can be factors.
- The file can have any number of other meta data columns; however, _they must all occur in the data frame **before** the columns of genetic data_.
- When you pass a file into any of these functions, you have to tell it which column the genetic data starts
in, and it is assumed that all the columns after that one contain genetic data.
- If you are doing a mixture analyis, the data frame of mixture fish and of the reference fish must have the
same column structure.


## An example reference data file

Here are the meta data columns and the first two loci for eight individuals in the `chinook` reference data set that comes with the 
package:
```{r}
library(rubias)
head(chinook[, 1:8])
```

## An example mixture data file

Here is the same for the mixture data frame that goes along with that reference data set:
```{r}
head(chinook_mix[, 1:8])
```


# Performing a Genetic Mixture Analysis

Not written.

# Simulation Assessment of Genetic References

If you want to know how much accuracy you can expect given a set of genetic markers and
a grouping of populations (`collection`s) into reporting units (`repunit`s), there are two
different functions you might use:

1. `assess_reference_loo()`: This function carries out simulation of mixtures using the leave-one-out
approach of @Andersonetal2008.
2. `assess_reference_mccv()`: This functions breaks the reference data set into different subsets, one of which
is used as the reference data set and the other the mixture.  It is difficult to simulate very large mixture 
samples using this method, because it is constrained by the number of fish in the reference data set.  
Additionally, there are constraints on the mixing proportions that can be simulated because of variation in the 
number of fish from each collection in the reference.

Both of the functions take two required arguments: 1) a data frame of reference genetic data, and 2) the 
number of the column in which the genetic data start.

## Specifying mixture proportions in `assess_reference_loo()`

By default, each iteration, the proportions of fish from each reporting unit is simulated
from a Dirichlet distribution with parameter (1.5,...,1.5).  And, within each reporting unit the mixing 
proportions from different collections are
drawn from a Dirichlet distribution with parameter  (1.5,...,1.5).

The value of 1.5 for the Dirichlet parameter for reporting units can be changed using the
`alpha_repunit`. The Dirichlet parameter for collections can be set using the `alpha_collection` parameter.

Sometimes, however, more control over the composition of the simulated mixtures is desired. This is achieved
by passing a two-column _data.frame_ to either `alpha_repunit` or `alpha_collection` (or both).  If you are 
passing the data.frame in for `alpha_repunit`, the first column must be named `repunit` and it must contain 
characters specifying reporting units.  In the data.frame for `alpha_collection` the first column must be 
named `collection` and must hold strings specifying different collections.  It is an error if
a repunit or collection is specified that does not exist in the reference.  However, you do not need to 
specify a value for every reporting unit or collection.  (If they are absent, the value is assumed to be zero.)

The second column of the data frame must be one of `count`, `ppn` or `dirichlet`.  These specify, respectively,
1) the exact count of individuals to be simulated from each repunit (or collection); 2) the proportion
of individuals from each repunit (or collection); or 3) the parameters of a Dirichlet distribution from which
the proportion of individuals should be simulated.  These `ppn` values will be normalized to sum to one if 
they do not.  As such, they can be regarded as weights.










# References
