---
title: "A Chinook-Like Data Set for Perfect Assignment"
output:
  html_notebook:
    toc: yes
  pdf_document:
    toc: yes
---


```{r setup, include=FALSE}
# set the working directory always to the project directory (one level up)
knitr::opts_knit$set(root.dir = normalizePath(rprojroot::find_rstudio_root_file())) 
```

libraries:
```{r load-libs}
library(tidyverse)
library(stringr)
library(rubias)
```


# Introduction and Making the Data set

In order to track down possible bugs and collection-ordering issues, I am going to
make a data set that is just like the `chinook` data set, but in which assignment to
different collections will be essentially perfect, by dint of each population being fixed
for a certain genetic signature.

We don't want to have a whole lot of loci, so we will use as few as we can: 
$\lceil\log_2(C)\rceil$, of them, where $C$ is the number of collections.  Basically we
will turn the index of each collection into a bitstring that tells us which loci are 
fixed for the alternate vs the reference allele (just 1 and 2, say), and then we will 
write that out to a file and read it back in.  Like so:

## A function to turn an integer into bits
We only want to keep the
first `d` bits...
```{r}
#' turn i into a string of d bits
int_to_bit_vec  <- function(i, d = 7) {
  as.character(intToBits(i))[1:d] %>%
    str_sub(., 2, 2) %>%
    as.integer()
}
```

Now, we will build on that to get something that turns a vector of 
indexes to a vector of strings of diploid genotypes.
```{r}
perf_genos <- function(v) {
  lapply(v, function(x) {
    paste( rep(int_to_bit_vec(x) + 1, each = 2), collapse = "\t")
  }) %>% 
    unlist
}
```
## Making genos for everyone

```{r}
binary_chinook <- chinook %>%
  mutate(coll_idx = as.integer(factor(collection, levels = unique(collection)))) %>%
  mutate(genostr = perf_genos(coll_idx)) %>%
  select(-coll_idx) %>%
  select(sample_type, repunit, collection, indiv, genostr)
```

Then we write this out to a file briefly and read it back in
```{r}
names(binary_chinook)[5] <- paste("loc", rep(1:7, each = 2), sep = "_", collapse = "\t")
tt <- tempfile()
write.table(binary_chinook, file = tt, quote = FALSE, sep = "\t", row.names = FALSE, col.names = TRUE)
```

Then read that back in
```{r}
pc <- read_tsv(tt)
```
That (`pc`) is what we have saved as `perfect_chinook` in the package.  


# Now, make a mixture in the proportions that we see in chinook_mix

I am going to make something that looks like `chinook_mix` by merely 
sampling individuals out of the reference and renaming them.

Here are the numbers that we want:
```{r}
desired_nums <- tibble(
  repunit = c("CentralValleyfa","CentralValleyfa","CentralValleysp","CentralValleyfa","KlamathR","RogueR","CentralValleyfa","CaliforniaCoast","CaliforniaCoast","NCaliforniaSOregonCoast","RogueR","KlamathR","UColumbiaRsufa","SnakeRfa","MidOregonCoast"),
  collection = c("Battle_Cr","Mokelumne_R_fa","Deer_Cr_sp","Feather_H_sp","Klamath_IGH_fa","Applegate_Cr","Sacramento_R_lf","Eel_R","Russian_R","Smith_R","Cole_Rivers_H","Trinity_H_sp","Hanford_Reach","Lyons_Ferry_H","Umpqua_sp"),
  num = c(321, 122, 84, 62, 46, 40, 20, 15, 9, 7, 6, 5, 3, 2, 2)
)
```

And here we make a template for putting all those in there:
```{r}
reppy <- 
  tibble(repunit = rep(desired_nums$repunit, desired_nums$num),
         collection = rep(desired_nums$collection, desired_nums$num)
         )
```
And now we just need to do some funky replicating to get all these 
genotypes
```{r}
pc_mix <- perfect_chinook %>% 
  group_by(repunit, collection) %>% 
  slice(1) %>% # this gets one individual from each collection
  left_join(reppy, .) %>%
  group_by(collection) %>%
  mutate(indiv = sprintf("fake_mix_%03d:%s", 1:n(), indiv)) %>%
  ungroup() %>% 
  mutate(sample_type = "mixture") %>%
  mutate(collection = "fake_perfect_mixture") %>%
  mutate(repunit = NA_character_) %>%
  select(sample_type, repunit, collection, indiv, everything())
```
Now, actually we want to make two mixture collections of this
```{r}
pc_mix2 <- bind_rows(pc_mix,
          pc_mix %>% mutate(collection = "fake_perfect_mixture_2", indiv = paste0(indiv, "-A")))
```

And this is what we have saved as `perfect_chinook_mix`.  


## Check infer_mixture()

Let's see how this is working.  And let's throw in a curve-ball by permuting the
rows of each data set around so that we don't have blocks of individuals from the same
reporting units and collections.
```{r}
set.seed(5)
pc_perm <- perfect_chinook[sample(1:nrow(perfect_chinook)), ]
pcm_perm <- perfect_chinook_mix[sample(1:nrow(perfect_chinook_mix)), ]
IM <- infer_mixture(pc_perm, pcm_perm, 5, method = "MCMC")
```

Now, compare this to what we know to be correct.
```{r}
true_nums <- perfect_chinook_mix %>%
  mutate(true_pop = str_replace_all(indiv, "fake_mix_", "") %>% str_replace_all("[0-9:]", "") ) %>%
  count(collection, true_pop) %>%
  rename(mixture_collection = collection, collection = true_pop) %>%
  arrange(mixture_collection, desc(n)) %>%
  mutate(collection = str_replace(collection, "-A$", "")) %>%
  left_join(., perfect_chinook %>% count(repunit, collection) %>% select(-n)) %>%
  select(mixture_collection, collection, repunit, n) %>%
  rename(true_num = n)
  

true_nums
```

First, just compare the collection proportions
```{r}
left_join(true_nums, IM$mixing_proportions) %>%
  group_by(mixture_collection) %>%
  mutate(true_pi = true_num / sum(true_num))
```

Yep, that is correct.

Now, let's check the indiv_posteriors.
```{r}
IM$indiv_posteriors %>%
  group_by(mixture_collection, indiv) %>%
  top_n(n = 1, wt = PofZ) %>%
  ungroup() %>%
  mutate(true_pop = str_replace_all(indiv, "fake_mix_", "") %>% str_replace_all("[0-9:]", "") ) %>%
  mutate(true_pop = str_replace(true_pop, "-A$", "")) %>%
  mutate(correct = true_pop == collection) %>%
  summarise(mean(correct))
```
Yep.  That check out.


## Check infer_mixture with the PB option

I was getting some funky results from this previously.  Let's see how it is working now.
First, let's not throw any permuted-rows curveballs...
```{r}
IM_pb <- infer_mixture(perfect_chinook, perfect_chinook_mix, 5, method = "PB")
```

At the end of that we can compare the regular mixing proportion estimates (which we know to
be correct) to the bootstrap-corrected ones.

```{r}
IM_pb$mixing_proportions %>%
  group_by(mixture_collection, repunit) %>%
  summarise(repunit_ppn = sum(pi)) %>%
  arrange(mixture_collection, desc(repunit_ppn)) %>%
  left_join(IM_pb$bootstrapped_proportions)
```

OK, that is clearly broken.  It appears that the problem is that the indexing of collections is not
correct in when the repunit proportions are being computed.  I think it is fine when all the collections
are grouped together within reporting units in the reference file, but we cannot assume that.  In the chinook
data the Feather_H_sp is actually in the Fall Run repunit and that seems to throw everything off.

