---
title: "Mixed Fishery Simulations -- Labrador"
output: 
  html_notebook:
    toc: true
    toc_float: true
---


Brendan had been trying this with `gsi_sim` and got some bugs.  In looking at the error message
I suspect it has something to do with a locus being missing in all the individuals in a population,
but I haven't looked into it too much.  Rather than get back into the guts of `gsi_sim` I thought
it would make better sense for me to just do the simulations using [rubias](https://github.com/eriqande/rubias),
because that is the software that I plan to support in the future (and because the interface is way easier
than gsi_sim---it is all R-based and largely tidyverse compliant!).

Brendan, if you want to use rubias (I recommend it over gsi_sim) go to the GitHub page and read the README.

So, let's do that. Load up some packages first.
```{r}
if (!("rubias" %in%  rownames(installed.packages()))) {
  devtools::install_github("eriqande/rubias")
}
library(tidyverse)
library(rubias)
```

## Converting gsi_sim data to rubias

There is a function for this in `rubias`.  

It looks like each population is a reporting unit in 
Brendan's file, so we will just do it that way.  I will do this on my system
but I don't expect it to work on Windows...(see `?read_gsi_sim`).

After we get the file, we set reporting units to the same as collections, and then we write
it to a file to be used later (which I will send to Brendan).
```{r, eval=FALSE}
labra <- read_gsi_sim("~/Downloads/All_Lab_Rivers_GSISIM.txt", sample_type = "reference") %>%
  mutate(repunit = collection)

# now, it looks like missing data are denoted by "000".  Let's turn those into NAs
labra[labra == "000"] <- NA

saveRDS(labra, file = "../data/all-lab-rivers-rubias.rds", compress = "xz")
```

So, now we can read that in again (this Brendan can do with to the file I send him...)
```{r}
labra <- readRDS("../data/all-lab-rivers-rubias.rds")
```
Look at a few rows to see that it is essentially a two column format:
```{r}
labra[1:10, 1:10]
```

## Self assignment exercise

First things first, let's just do self-assignment of individuals. Note that this function gives a little summary of the data
so you can make sure that it is working correctly!
```{r}
sa_labra <- self_assign(labra, 5)
```

Then we can have a look at how often individuals are assigned (useing maximum posterior) to their own population.
```{r}
map_ass <- sa_labra %>%
  group_by(indiv) %>%
  top_n(n = 1, wt = scaled_likelihood) %>%
  ungroup

self_ass_mat <- map_ass %>%
  count(collection, inferred_collection) %>%
  spread(key = inferred_collection, value = n, fill = 0)

self_ass_mat
```

Great! That is looking pretty good.

## Simulations of mixed fisheries

### Default settings

Let's do these first with samples of size 500 and just use the default 
parameter values which "spray proportions of fish around randomly."  Let's do
100 reps.  Also, by default, this does the CV-ML method from Anderson, Waples,
and Kalinowski (2008), which resamples full multilocus genotypes rather than
gene copies.  The following takes about 20 seconds or so.
```{r, message=FALSE}
loo_sim1 <- assess_reference_loo(labra, 5, reps = 100, mixsize = 500)
```
Then we can easily plot the results:
```{r, fig.width = 10, fig.height = 14}
ggplot(loo_sim1, aes(x = true_pi, y = post_mean_pi, colour = collection)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  facet_wrap(~collection, ncol = 4)
```

That is revealing.  Everything looks pretty good but POT is a disaster (saw that in the self-assignment results, too).
The POT fish probably are ending up in MIDBTLUWST, and there are some mild biases in some other pops (SAN a little
funky).  But overall, this looks like a pretty great set of markers.

Let's see what it looks like if we resample gene-copies rather than individuals:
```{r, message=FALSE}
loo_sim2 <- assess_reference_loo(labra, 5, reps = 100, mixsize = 500, resampling_unit = "gene_copies")
```
Then we can easily plot the results:
```{r, fig.width = 10, fig.height = 14}
ggplot(loo_sim2, aes(x = true_pi, y = post_mean_pi, colour = collection)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  facet_wrap(~collection, ncol = 4)
```
Those results are pretty comparable.  Which is good.

### 100 from each collection

If you wanted to simulate exactly 100 from each collection (like you had in the mix1000 comms file)
you would do like this: first make a data frame of the desired collection counts
```{r}
coll_counts <- labra %>%
  count(collection) %>%
  select(-n) %>%
  mutate(cnt = 100)

coll_counts

```
Then, pass that in for `alpha_collection`.  Let's do 20 reps.
```{r}
loo_sim3 <- assess_reference_loo(labra, 5, reps = 20, mixsize = 500, alpha_collection = coll_counts)
```
Then again, plot the result just for fun, to see that all the true values are at 100.
```{r, fig.width = 10, fig.height = 14}
ggplot(loo_sim3, aes(x = n, y = post_mean_pi, colour = collection)) +
  geom_point() +
  facet_wrap(~collection, ncol = 4) +
  geom_hline(yintercept = 1.0 / nrow(coll_counts))
```
We put a horizontal line at the true value for each population.

## Wrap up

That is it for now.  `rubias` is way easier to use than gsi_sim.  It also can analyze multiple 
actual mixture samples in one fell swoop.  See `infer_mixture`. 

