---
title: "Known Collections on DFO Real Data"
output: html_notebook
---

Colin sent me some real data to play with.  Very cool.

I have put it into development/private_data and have gitignored it all so it won't
inadvertently get put in the cloud somewhere...

Let's read it in and see what happens.

```{r}
library(tidyverse)
library(rubias)

bco <- read_tsv("../private_data/mco_escapes_rubias_Feb_12_2018/bco_rubias_Feb_28_2018.txt.gz")
mco <- read_tsv("../private_data/mco_escapes_rubias_Feb_12_2018/mco_escapes_rubias_Feb_12_2018.txt")

```


Let's see what we have for known_collection individuals in the mixture:
```{r}
kc_counts <- count(mco, collection, known_collection)
kc_counts
```

Note, for future reference, that Rosewall_Cr fish is 105_2017_54908_F.

OK. We have 5 mixture collections, which are probably individuals escaping to each of the rivers there.

The known collection ones are resolved using PBT.

Let us very quickly compare the results that we get for each of the known_collections 
occurring in mco (there are 6 of them), when we either use the known_collection information, or we don't.

We will focus on collections rather than reporting units, since they should be harder to
resolve, and therefore there will be more difference expected when you run it with vs without
the known collections.

## Run the MCMC, etc

Here we do it with the known collections
```{r}
wkc <- infer_mixture(bco, mco, 6)
```


Here we do it without the known_collections
```{r}
nkc <-infer_mixture(bco[-1], mco[-1], 5)
```

### Messing with pi_prior_pseudo_count_sum
Here we do it both with and without, but we set the pi_prior_pseudo_count_sum to be 0.05.

```{r}
wkc_pp <- infer_mixture(bco, mco, 6, pi_prior_pseudo_count_sum = 0.05)
nkc_pp <-infer_mixture(bco[-1], mco[-1], 5, pi_prior_pseudo_count_sum = 0.05)
```


Merge together the mix_prop_traces for each of those so we can analyse it together.
```{r}
mer <- bind_rows(list(with_kc = wkc$mix_prop_traces, 
                      without_kc = nkc$mix_prop_traces,
                      with_kc_pp05 = wkc_pp$mix_prop_traces, 
                      without_kc_pp05 = nkc_pp$mix_prop_traces),
                 .id = "how_run")
```

Now pick out the populations we are particularly interested in:
```{r}
tmp <- kc_counts %>%
  filter(!is.na(known_collection)) %>%
  select(-n) %>%
  rename(mixture_collection = collection) %>%
  rename(collection = known_collection)

slim <- mer %>%
  semi_join(., tmp, by = c("mixture_collection", "collection")) %>%
  filter(sweep > 500)  # burn-in

slim %>% group_by(mixture_collection, collection, how_run) %>% summarise(mean = mean(pi), sd = sd(pi)) 
```

And what we see here is very little difference. Which is what we expect.  The one Rosewall_Cr fish identified by PBT
makes the estimated fraction of Rosewall_Cr in the Puntledge_R mixture collection much higher when using the known
collections, and, as a consequence, the sd is higher, but that is just a consequence of the fact that sd will be higher if the mean is higher.  

The high confidence with which individuals are assigned using GSI can be seen by looking at the
individual assignments.  Run this if you want to peruse them:
```{r, eval=FALSE}
View(nkc$indiv_posteriors %>% group_by(mixture_collection, indiv) %>% filter(near(PofZ, max(PofZ))))
```

So, overall, it looks like the known_collection option is working the way that it should, but there isn't a
noticeable difference because the resolution with GSI is already quite good, and most of these fish are
in mixtures where most of the fish are from one location.  

Does that seem reasonable?  Please let me know if not.

