---
title: "Helping Kerry With Stuff"
output: html_notebook
---


First get data and load libs
```{r}
library(tidyverse)
library(stringr)
library(rubias)

ale_ref <- read_csv("../data/alewife-baseline-input_02062017.csv.gz")
ale_mix <- read_csv("../data/alewife-bycatch-input_02062017.csv.gz")
```


Let's toss out collections that don't have at least 20 indivs
```{r}
ale_ref2 <- ale_ref %>%
  group_by(repunit, collection) %>%
  filter(n() > 20) %>%
  ungroup()
```
## Assessing the reference

Let's see how we do this.  First using leave one out type procedure:
```{r}
arl <- assess_reference_loo(reference = ale_ref2, gen_start_col = 5)


```

Then plot the result...
First by population
```{r, fig.width = 15, fig.height=20}
ggplot(arl, aes(x = omega, y = post_mean, colour = repunit)) +
  geom_point() +
  facet_wrap(~ collection)
         
```

And then aggregate by repunit:
```{r}
arl_ru <- arl %>%
  group_by(iter, repunit) %>% 
  summarise(true_repu_prop = sum(omega), est_repu_prop = sum(post_mean), true_repu_num = sum(n)) %>%
  ungroup()
```
And plot that:
```{r}
ggplot(arl_ru, aes(x = true_repu_prop, y = est_repu_prop, colour = repunit)) +
  geom_point() +
  facet_wrap(~ repunit) +
  geom_abline(intercept = 0, slope = 1)
  
```


## Estimate the mixing proportions of the bycatch against the baseline

```{r}
mix_est <- infer_mixture(reference = ale_ref2, mixture = ale_mix, gen_start_col = 5)
```


The result is a list.  If we want proportions we want `mix_est$mixing_proportions`

Once again we have to aggregate it over reporting units:
```{r}
ru_mix_ests <- mix_est$mixing_proportions %>% 
  group_by(mixture_collection, repunit) %>%
  summarise(repu_prop = sum(pi))
```

Now make a bar plot:
```{r}
ggplot(ru_mix_ests, aes(x = mixture_collection, y = repu_prop, fill = repunit)) + 
  geom_col() +
  coord_flip()
```


Let's see what happens if we do the parametric bootstrapping (just with the repunits as they are)
```{r}
mix_est_pb <- infer_mixture(reference = ale_ref2, mixture = ale_mix, gen_start_col = 5, method = "PB")
```
