---
title: "Alewife SNP Baselines"
output:
  html_notebook:
    toc: yes
  pdf_document:
    toc: yes
---


```{r setup, include=FALSE}
# set the working directory always to the project directory (one level up)
knitr::opts_knit$set(root.dir = normalizePath(rprojroot::find_rstudio_root_file())) 
```

libraries:
```{r load-libs}
library(tidyverse)
library(rubias)
library(stringr)
```

## Get The Data

First I need to read the data in and format it the way we need it.
```{r proc-data}
genos <- read_csv("development/data/snp-alewife-stuff/alewife-baseline.csv.gz") %>%
  select(-(`Pop order`:`Modified IND`), -(`Drainage by year code`:Year)) %>%
  rename(indiv = `Drainage code`) %>%
  tidyr::separate(data = ., col = indiv, into = c("collection", "idnum"), remove = FALSE) %>%
  select(-idnum) %>%
  mutate(sample_type = "reference") %>%
  select(sample_type, collection, indiv, everything())
  
repu <- read_tsv("development/data/snp-alewife-stuff/alewife_7_repgroups.txt") %>%
  rename(collection = Pop,
         repunit = RepGroup) %>%
  select(-RepGroupNum)

ale_snps7 <- left_join(genos, repu) %>%
  select(sample_type, repunit, everything())
  
```

## Simulate and assess reference

OK, now let us do the standard simulation on these guys.
```{r sim-and-assess}
saar7 <- assess_reference_loo(reference = ale_snps7, 
                                       gen_start_col = 5, 
                                       reps = 100, 
                                       mixsize = 200)
```

summarize that to repunit:
```{r sum-to-rep}
rep7 <- saar7 %>%
  group_by(iter, repunit) %>%
  summarize(post_mean = sum(post_mean),
            mle = sum(mle),
            true_rep_prop = sum(omega),
            n = sum(n)) %>%
  group_by(iter) %>%
  mutate(samp_frac = n / sum(n))
```
plot it.

```{r plot7, fig.width=7}
ggplot(rep7, aes(x = true_rep_prop, y = post_mean, colour = repunit)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  facet_wrap(~ repunit) 
```
OK, that is cool.  MB, NUN, BIS, and LIS are four groups that would traditionally all
be lumped together as "SNE".  There are problems with them, clearly, but to be honest
they don't look all that bad.  I am curious to see if the parametric bootstrap
bias correction might help things.

Let's also look at the sample proportion, which will lessen the
variance due to just random sampling...
```{r plot7-2, fig.width=7}
ggplot(rep7, aes(x = samp_frac, y = post_mean, colour = repunit)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  facet_wrap(~ repunit) 
```

## Bias comparison function

Let's just give this a whirl.  There is a problem, however: collection MED has 2 indivs
and NAN has only 9.  This causes problems for `mixture_draw()` it appears.  So, let's just 
eliminate those to collections:

```{r bias-comp, eval=FALSE}
as7_filt <- ale_snps7 %>% 
  filter(!(collection %in% c("MED", "NAN"))) %>%
  mutate(collection = factor(collection, levels = unique(collection)))


bc <- assess_pb_bias_correction(reference = as7_filt, gen_start_col = 5, seed = 5, mixsize = 200, nreps = 100)
```

2 or 3 hours later we have the results we want.  I saved those results and we can read them 
back in:
```{r read-output}
bc <- readRDS("development/outputs/alewife_snps_100_of_200.rds")
```

Then plot it.
```{r plot, fig.width = 8, fig.height = 10}
#saveRDS(bc, "alewife_snps_100_of_200.rds")
rho_data <- bc %>%
      tidyr::gather(key = "method", value = "rho_est", rho_mcmc:rho_pb)

g <- ggplot2::ggplot(rho_data, ggplot2::aes(x = true_rho, y = rho_est, colour = repunit)) +
      ggplot2::geom_point() +
      ggplot2::facet_grid(repunit ~ method) +
      ggplot2::geom_abline(intercept = 0, slope = 1)

g
```
