---
title: "Testing assess_reference_mc"
author: "benmo"
date: "June 20, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

libraries:
```{r load-libs}
library(tidyverse)
library(stringr)
library(rubias)
```

## Introduction

This document will build on the `perfect_chinook` data set to test the tolerance
of `assess_reference_mc` for datasets other than `alewife`. Namely, we are
interested in whether the function works properly when reporting units and collections
are not present within continuous blocks of rows in the dataset.

So, we begin by trying a simple `assess_reference_mc` on the `perfect_chinook` dataset.

```{r}
set.seed(15)
out_unperm <- assess_reference_mc(perfect_chinook, 5)
```

First rows look good, but we graph to see how it holds across the range of possible rhos.

```{r}
rhos_unperm <- out_unperm %>%
  group_by(iter,repunit) %>%
  summarise(true_rho = sum(omega), rho_est = sum(post_mean))

rg <- ggplot2::ggplot(rhos_unperm, ggplot2::aes(x = true_rho, y = rho_est, colour = repunit)) +
      ggplot2::geom_point() +
      ggplot2::facet_wrap(~repunit) +
      ggplot2::geom_abline(intercept = 0, slope = 1)
print(rg)
```

Printing `rg` from the console shows that everything is looking good.

We then permute the order of individuals in `perfect_chinook`, as before:

```{r}
set.seed(5)
pc_perm <- perfect_chinook[sample(1:nrow(perfect_chinook)), ]
out_perm <- assess_reference_mc(pc_perm, 5)
```

First set looks good. Let's summarise rho and graph.

```{r}
rhos_perm <- out_perm %>%
  group_by(iter,repunit) %>%
  summarise(true_rho = sum(omega), rho_est = sum(post_mean))

rgp <- ggplot2::ggplot(rhos_perm, ggplot2::aes(x = true_rho, y = rho_est, colour = repunit)) +
      ggplot2::geom_point() +
      ggplot2::facet_wrap(~repunit) +
      ggplot2::geom_abline(intercept = 0, slope = 1)
print(rgp)
```

It looks just as good as the other. Therefore, it seems clear that `assess_reference_mc`
handles the reporting unit & collection sorting appropriately.

The Central Valley Fall estimate is noticeably biased downwards at small rho values. 
This doesn't line up with prior cases of RU bias; it's below the $N_{c}/P$ value of $8/69$,
so if anything one would expect it to be biased upwards. 
