---
title: "Testing the `simulate_random_samples()` function"
output: 
  html_notebook:
    toc: true
---


```{r setup, include=FALSE}
# set the working directory always to the project directory (one level up)
knitr::opts_knit$set(root.dir = normalizePath(rprojroot::find_rstudio_root_file())) 
```

libraries:
```{r load-libs}
library(tidyverse)
library(rubias)
library(stringr)
```

My goal here is to test the many different scenarios by which the composition of simulated
samples can be specified for doing simulations with rubias.

I just overhauled the `simulate_random_samples()` function and I want to both document its
use and test it.  

Most users will never call it directly, but I will here to make sure that things are working
correctly.


# Preliminaries

We will do this with the `chinook` data set.  Here is what the repunits and collections look like
in that data set
```{r}
repc <- chinook %>%
  group_by(repunit, collection) %>%
  tally() %>%
  select(-n)

repc
```

We will use that to left_join results on later.  

In order to call `simulated_random_samples()` we need to get some variables, that we get
this way:
```{r}
p <- tcf2param_list(chinook, 5, summ = F)
```


## A function to extract simulated values

Given a list of multiple outputs of simulate_random_samples() (obtained by lapplying
multiple reps for example), this function collects the output into three tidy data frames.
```{r}
tidy_srs <- function(L, repc) {
  # get the rho values
  rho <- lapply(L, function(x) {
    enframe(x$rho, name = "repunit", value = "rho")
    }) %>% 
    bind_rows(.id = "rep") %>%
    mutate(rep = as.integer(rep))
  
  # and then get the omega and n's
  omegan <- lapply(L, function(x) {
    ns <- as.data.frame(table(names(x$sim_coll))) %>% 
      rename(collection = Var1, n = Freq) %>%
      mutate(collection = as.character(collection)) %>%
      as_tibble()
    
    enframe(x$omega, name = "collection", value = "omega") %>%
      left_join(., ns, by = "collection")
    }) %>% 
    bind_rows(.id = "rep") %>%
    mutate(rep = as.integer(rep)) %>%
    mutate(n = ifelse(is.na(n), 0, n)) %>%
    left_join(repc, ., by = "collection")
  
  list(rho = rho, omegan = omegan)
}

### Here is a function for summarizing those and returning those summarized values in a list
# along with the originals.
summ_it <- function(S) {
  rho_summ <- S$rho %>%
    group_by(repunit) %>%
    summarise(rho_mean = mean(rho),
              rho_var = var(rho),
              rho_min = min(rho),
              rho_max = max(rho))
  
  omegan_summ <- S$omegan %>%
    group_by(repunit, collection) %>%
    summarise(omega_mean = mean(omega),
              n_mean = mean(n),
              omega_var = var(omega),
              n_var = var(n),
              omega_min = min(omega),
              omega_max = max(omega),
              n_min = min(n),
              n_max = max(n)) %>%
    group_by(repunit) %>%
    mutate(summed_rho = sum(omega_mean),
           npops = n())
  
  S$rho_summ <- rho_summ
  S$omegan_summ <- omegan_summ
  
  S
}
```

In order to put things in a good order on plots we make this:
```{r}
repcF <- repc %>%
  mutate(repu_coll = paste(repunit, collection, sep = "--")) %>%
  select(repu_coll, everything())

repcF$repu_coll <- factor(repcF$repu_coll, levels = unique(repcF$repu_coll))
```

# repunit single

First gonna just test the repunit "single" value setting.

## collection single

This is the simplest, default case.  Let's get all the output:
```{r}
res <- lapply(1:1000, function(x) {
  simulate_random_samples(RU_starts = p$RU_starts, RU_vec = p$RU_vec, size = 100)
  }) %>%
  tidy_srs(., repc) %>%
  summ_it()
```
Then look at it:

### rho values
```{r}
ggplot(res$rho_summ, aes(x = repunit, y = rho_mean)) +
  geom_point() +
  geom_segment(aes(y = rho_min, yend = rho_max, xend = repunit)) +
  coord_flip()
```



### omega values

```{r}
ggplot(res$omegan_summ, aes(x = collection, y = omega_mean, colour = as.factor(npops))) +
  geom_point() +
  geom_segment(aes(y = omega_min, yend = omega_max, xend = collection)) +
  coord_flip()
```

Finally look at the number of simulated individuals:
```{r}
res$omegan_summ %>% 
  group_by(repunit) %>% 
  summarise(repu_n = sum(n_mean))
```
That checks out summing over reporting units.

And how about the number of collection, scaled by the number of collections in the repunit?
```{r}
res$omegan_summ %>%
  mutate(scaled_n = n_mean * npops) %>%
  select(repunit, collection, scaled_n, everything())
```

I had introduced a bug, but have fixed it now.  It seems to me that the omega vector that 
Ben carried around in `simulate_random_samples()` must have been ordered differently than the
natural `levels(collections)` order.  When I ran the above code (with a little `[1,]` thrown in for
the `rho` values, with the older version of the code that Ben wrote, the # of fish from each collection
is correct, but the omegas are wrong, suggesting that they are not ordered they way I thought they were.

When I sample the sim_colls I now do so from a a vector that
is ordered by the levels of the collections and it seems to work.



## collection sub_dirichlet

For this rather contrived case, the reporting unit proportions are drawn from a Dirichlet(1.5), but 
then I can specify different proportions within rep_units of the collections, following a dirichlet
distribution.  For this test, everyone will get the default 1.5 except for a few collections in
the CV fall and CV spring and California Coast.  Namely, in my examples, I have:
```{r}
sim_spec_examples$coll_sub_dirichlet
```

Let's give this thing a whirl.  Just like before:


```{r}
csd <- lapply(1:1000, function(x) {
  simulate_random_samples(RU_starts = p$RU_starts, RU_vec = p$RU_vec, size = 100,
                          alpha_collection = sim_spec_examples$coll_sub_dirichlet)}) %>% 
  tidy_srs(., repc) %>%
  summ_it()
```

### rho values

```{r}
ggplot(csd$rho_summ, aes(x = repunit, y = rho_mean)) +
  geom_point() +
  geom_segment(aes(y = rho_min, yend = rho_max, xend = repunit)) +
  coord_flip()
```

Those look correct!

### omega values
  Now, how about the omegas?
```{r, fig.height = 5}
ggplot(left_join(repcF, csd$omegan_summ), aes(x = repu_coll, y = omega_mean, colour = as.factor(npops))) +
  geom_point() +
  geom_segment(aes(y = omega_min, yend = omega_max, xend = repu_coll)) +
  coord_flip()
```

OK, that checks out.  Now check sim_colls
```{r}
csd$omegan_summ %>%
  left_join(repcF, .) %>%
  select(repu_coll, omega_mean, n_mean)
```

That check outs.  Good!

## collection sub_ppn

We are going to do something similar, but we will specify weights for the collections, and in 
this case, the default is practically 0.  So, for reporting units in which we specify at least one
collection with a weight, they will get pretty much all of it.
```{r}
sim_spec_examples$coll_sub_ppn
```

```{r}
csp <- lapply(1:1000, function(x) {
  simulate_random_samples(RU_starts = p$RU_starts, RU_vec = p$RU_vec, size = 100,
                          alpha_collection = sim_spec_examples$coll_sub_ppn)}) %>% 
  tidy_srs(., repc) %>%
  summ_it()
```

### rho values

```{r}
ggplot(csp$rho_summ, aes(x = repunit, y = rho_mean)) +
  geom_point() +
  geom_segment(aes(y = rho_min, yend = rho_max, xend = repunit)) +
  coord_flip()
```

Those look correct!

### omega values
  Now, how about the omegas?
```{r, fig.height = 5}
ggplot(left_join(repcF, csp$omegan_summ), aes(x = repu_coll, y = omega_mean, colour = as.factor(npops))) +
  geom_point() +
  geom_segment(aes(y = omega_min, yend = omega_max, xend = repu_coll)) +
  coord_flip()
```


OK, that checks out.  Now check sim_colls
```{r}
csp$omegan_summ %>%
  left_join(repcF, .) %>%
  select(repu_coll, omega_mean, n_mean)
```

Yep. That checks out.



## Collection Dirichlet

This is best done whilst not setting anything for alpha_repunit (else you get lots of warnings).
We are going to make a contrived example here in which each reporting unit gets a certain weight, and that
gets split between the collections in it according to their position in the reporting unit. This
is how I did it, and it is now part of the package.
```{r eval = FALSE}
set.seed(555)
ppn <- repc %>%
  group_by(repunit) %>%
  mutate(repwt = rep(runif(1, min = 1, max = 100), n())) %>%
  mutate(ppn = repwt / 1:n()) %>%
  ungroup() %>%
  select(collection, ppn)

sim_spec_examples$coll_ppn <- ppn


sim_spec_examples$coll_dirichlet <- ppn %>%
  rename(dirichlet = ppn)
  
```

Here is what the input looks like:
```{r}
sim_spec_examples$coll_dirichlet
```
So, let's run it:

```{r}
cdir <- lapply(1:1000, function(x) {
  simulate_random_samples(RU_starts = p$RU_starts, RU_vec = p$RU_vec, size = 100,
                          alpha_collection = sim_spec_examples$coll_dirichlet)}) %>% 
  tidy_srs(., repc) %>%
  summ_it()
```

Then join it to what we asked for:
```{r}
tmp <- sim_spec_examples$coll_dirichlet %>%
  mutate(n_exp = 100 * dirichlet / sum(dirichlet),
         omega_exp = dirichlet / sum(dirichlet))
comp_cdir <- left_join(tmp, cdir$omegan_summ)
comp_cdir
```

Booyah!  That looks just right:
```{r}
ggplot(comp_cdir, aes(x = n_exp, y = n_mean)) + 
  geom_point() +
  geom_point(aes(x = omega_exp * 100, y = omega_mean * 100), colour = "red")
```


## Collection Ppn

We are going to do the same, but have them interpreted as collection proportions, rather
than dirichlet parameters:
```{r}
sim_spec_examples$coll_ppn
```

And here it goes:
```{r}
cppn <- lapply(1:1000, function(x) {
  simulate_random_samples(RU_starts = p$RU_starts, RU_vec = p$RU_vec, size = 100,
                          alpha_collection = sim_spec_examples$coll_ppn)}) %>% 
  tidy_srs(., repc) %>%
  summ_it()
```

```{r}
tmp <- sim_spec_examples$coll_ppn %>%
  mutate(n_exp = 100 * ppn / sum(ppn),
         omega_exp = ppn / sum(ppn))
comp_cppn <- left_join(tmp, cppn$omegan_summ)
comp_cppn
```
Yep.  That is all good. Note that the variance of omega is 0 there.

## Collection cnt

Now, let's see if we can simulate a lot of samples with exactly the same number
of individuals each time.
```{r}
ccnt <- lapply(1:1000, function(x) {
  simulate_random_samples(RU_starts = p$RU_starts, RU_vec = p$RU_vec, size = 100,
                          alpha_collection = sim_spec_examples$coll_cnt)}) %>% 
  tidy_srs(., repc) %>%
  summ_it()
```
OK, we get that warning.  That is cool.  In the future I think I could
throttle that warning and just say that `size` is ignored in the docs.

```{r}
tmp <- sim_spec_examples$coll_cnt %>%
  mutate(omega_exp = cnt / sum(cnt))
comp_ccnt <- left_join(tmp, ccnt$omegan_summ)
comp_ccnt
```

Yep, that all checks out.

# repunit dirichlet

So, we need to test this with the three different "sub_collection"s.

First we make up a set of Dirichlet parameters for the different repunits
```{r, eval = FALSE}
set.seed(555)
sim_spec_examples$repunit_dirichlet <- tibble(
  repunit = unique(repc$repunit)) %>% 
  mutate(dirichlet = runif(n(), min = -15, max = 15)) %>%
  filter(dirichlet > 0)
```
## collection single

First we test the situation where collection is just give the default dirichlet parameter:
```{r}
rdir <- lapply(1:1000, function(x) {
  simulate_random_samples(RU_starts = p$RU_starts, RU_vec = p$RU_vec, size = 100,
                          alpha_repunit = sim_spec_examples$repunit_dirichlet)}) %>% 
  tidy_srs(., repc) %>%
  summ_it()
```
Then we compare that to what we expected
```{r}
sim_spec_examples$repunit_dirichlet %>%
  mutate(rho_exp = dirichlet / sum(dirichlet)) %>%
  left_join(rdir$rho_summ, .)

```

And at the collection level:
```{r}
sim_spec_examples$repunit_dirichlet %>%
  mutate(rho_exp = dirichlet / sum(dirichlet)) %>%
  left_join(rdir$omegan_summ, .)
  
```

That all looks good.

## collection sub-dirichlet

```{r}
rdir_csd <- lapply(1:1000, function(x) {
  simulate_random_samples(RU_starts = p$RU_starts, RU_vec = p$RU_vec, size = 100,
                          alpha_repunit = sim_spec_examples$repunit_dirichlet,
                          alpha_collection = sim_spec_examples$coll_sub_dirichlet)}) %>% 
  tidy_srs(., repc) %>%
  summ_it()
```

Let's compute the expected proportions:
```{r}
sim_spec_examples$repunit_dirichlet %>%
  ungroup() %>%
  mutate(dirch_ppn = dirichlet / sum(dirichlet)) %>%
  left_join(repc, .) %>%
  left_join(., sim_spec_examples$coll_sub_dirichlet) %>%
  mutate(sub_dirichlet = ifelse(is.na(sub_dirichlet), 1.5, sub_dirichlet),
         dirichlet = ifelse(is.na(dirichlet), 1.5, dirichlet)) %>%
  group_by(repunit) %>%
  mutate(exp_omega = dirch_ppn * sub_dirichlet / sum(sub_dirichlet))  %>%
  left_join(., rdir_csd$omegan_summ)
```
That looks like it all checks out.

## collection sub_ppn

```{r}
rdir_csp <- lapply(1:1000, function(x) {
  simulate_random_samples(RU_starts = p$RU_starts, RU_vec = p$RU_vec, size = 100,
                          alpha_repunit = sim_spec_examples$repunit_dirichlet,
                          alpha_collection = sim_spec_examples$coll_sub_ppn)}) %>% 
  tidy_srs(., repc) %>%
  summ_it()
```

Let's compute the expected proportions:
```{r}
sim_spec_examples$repunit_dirichlet %>%
  ungroup() %>%
  mutate(dirch_ppn = dirichlet / sum(dirichlet)) %>%
  left_join(repc, .) %>%
  left_join(., sim_spec_examples$coll_sub_ppn) %>%
  mutate(sub_ppn = ifelse(is.na(sub_ppn), 0.0000001, sub_ppn),
         dirichlet = ifelse(is.na(dirichlet), 1.5, dirichlet)) %>%
  group_by(repunit) %>%
  mutate(exp_omega = dirch_ppn * sub_ppn / sum(sub_ppn))  %>%
  left_join(., rdir_csp$omegan_summ)
```
That looks like it checks out.

# repunit ppn

## collection single

First we test the situation where collection is just give the default dirichlet parameter:
```{r}
rppn_csing <- lapply(1:1000, function(x) {
  simulate_random_samples(RU_starts = p$RU_starts, RU_vec = p$RU_vec, size = 100,
                          alpha_repunit = sim_spec_examples$repunit_ppn)}) %>% 
  tidy_srs(., repc) %>%
  summ_it()
```
Then we compare that to what we expected
```{r}
sim_spec_examples$repunit_ppn %>%
  mutate(rho_exp = ppn / sum(ppn)) %>%
  left_join(rppn_csing$rho_summ, .)

```

And at the collection level:
```{r}
sim_spec_examples$repunit_ppn %>%
  mutate(rho_exp = ppn / sum(ppn)) %>%
  left_join(rppn_csing$omegan_summ, .)
  
```


## Collection sub_dirichlet

```{r}
rppn_csd <- lapply(1:1000, function(x) {
  simulate_random_samples(RU_starts = p$RU_starts, RU_vec = p$RU_vec, size = 100,
                          alpha_repunit = sim_spec_examples$repunit_ppn,
                          alpha_collection = sim_spec_examples$coll_sub_dirichlet)}) %>% 
  tidy_srs(., repc) %>%
  summ_it()
```

Let's compute the expected proportions:
```{r}
sim_spec_examples$repunit_ppn %>%
  ungroup() %>%
  mutate(scaled_ppn = ppn / sum(ppn)) %>%
  left_join(repc, .) %>%
  left_join(., sim_spec_examples$coll_sub_dirichlet) %>%
  mutate(sub_dirichlet = ifelse(is.na(sub_dirichlet), 1.5, sub_dirichlet),
         ppn = ifelse(is.na(ppn), 0, ppn)) %>%
  group_by(repunit) %>%
  mutate(exp_omega = scaled_ppn * sub_dirichlet / sum(sub_dirichlet))  %>%
  left_join(., rppn_csd$omegan_summ)
```
Looks like that works.



## Collection sub_ppn

```{r}
rppn_csp <- lapply(1:1000, function(x) {
  simulate_random_samples(RU_starts = p$RU_starts, RU_vec = p$RU_vec, size = 100,
                          alpha_repunit = sim_spec_examples$repunit_ppn,
                          alpha_collection = sim_spec_examples$coll_sub_ppn)}) %>% 
  tidy_srs(., repc) %>%
  summ_it()
```

Let's compute the expected proportions:
```{r}
sim_spec_examples$repunit_ppn %>%
  ungroup() %>%
  mutate(scaled_ppn = ppn / sum(ppn)) %>%
  left_join(repc, .) %>%
  left_join(., sim_spec_examples$coll_sub_ppn) %>%
  mutate(sub_ppn = ifelse(is.na(sub_ppn), 0.0000001, sub_ppn),
         ppn = ifelse(is.na(ppn), 0, ppn)) %>%
  group_by(repunit) %>%
  mutate(exp_omega = scaled_ppn * sub_ppn / sum(sub_ppn))  %>%
  left_join(., rppn_csp$omegan_summ)
```
Yep, those are right on.  And there is no variance in the omega's just like we expect.


# finally, repunit counts

All that remains is to do explicit counts in the repunits.

## collection single

First we test the situation where collection is just give the default dirichlet parameter:
```{r}
rcnt_csing <- lapply(1:1000, function(x) {
  simulate_random_samples(RU_starts = p$RU_starts, RU_vec = p$RU_vec, size = 100,
                          alpha_repunit = sim_spec_examples$repunit_cnt)}) %>% 
  tidy_srs(., repc) %>%
  summ_it()
```
Then we compare that to what we expected
```{r}
sim_spec_examples$repunit_cnt %>%
  mutate(rho_exp = cnt / sum(cnt)) %>%
  left_join(rppn_csing$rho_summ, .)

```
 Looks right.  
 
 
And at the collection level:
```{r}
sim_spec_examples$repunit_cnt %>%
  mutate(rho_exp = cnt / sum(cnt)) %>%
  left_join(rcnt_csing$omegan_summ, .)
  
```

That looks right.

## Collection sub_dirichlet

```{r}
rcnt_csd <- lapply(1:1000, function(x) {
  simulate_random_samples(RU_starts = p$RU_starts, RU_vec = p$RU_vec, size = 100,
                          alpha_repunit = sim_spec_examples$repunit_cnt,
                          alpha_collection = sim_spec_examples$coll_sub_dirichlet)}) %>% 
  tidy_srs(., repc) %>%
  summ_it()
```

Let's compute the expected proportions:
```{r}
sim_spec_examples$repunit_cnt %>%
  ungroup() %>%
  mutate(scaled_ppn = cnt / sum(cnt)) %>%
  left_join(repc, .) %>%
  left_join(., sim_spec_examples$coll_sub_dirichlet) %>%
  mutate(sub_dirichlet = ifelse(is.na(sub_dirichlet), 1.5, sub_dirichlet),
         cnt = ifelse(is.na(cnt), 0, cnt)) %>%
  group_by(repunit) %>%
  mutate(exp_omega = scaled_ppn * sub_dirichlet / sum(sub_dirichlet))  %>%
  left_join(., rppn_csd$omegan_summ)
```
Close, but not totally right on.  But I don't think I will worry too much about it.

```{r}
rcnt_csd$omegan_summ %>%
  group_by(repunit) %>%
  summarise(summed_n = sum(n_mean)) %>%
  left_join(., sim_spec_examples$repunit_cnt)
```
That part looks great.

## Collection sub_ppn

```{r}
rcnt_csp <- lapply(1:1000, function(x) {
  simulate_random_samples(RU_starts = p$RU_starts, RU_vec = p$RU_vec, size = 100,
                          alpha_repunit = sim_spec_examples$repunit_cnt,
                          alpha_collection = sim_spec_examples$coll_sub_ppn)}) %>% 
  tidy_srs(., repc) %>%
  summ_it()
```

Let's compute the expected proportions:
```{r}
sim_spec_examples$repunit_cnt %>%
  ungroup() %>%
  mutate(scaled_ppn = cnt / sum(cnt)) %>%
  left_join(repc, .) %>%
  left_join(., sim_spec_examples$coll_sub_ppn) %>%
  mutate(sub_ppn = ifelse(is.na(sub_ppn), 0.0000001, sub_ppn),
         cnt = ifelse(is.na(cnt), 0, cnt)) %>%
  group_by(repunit) %>%
  mutate(exp_omega = scaled_ppn * sub_ppn / sum(sub_ppn))  %>%
  left_join(., rcnt_csp$omegan_summ) %>%
  ungroup() %>%
  mutate(n_mean_ppn = n_mean / sum(n_mean))
```


Good.

# Finally, check for errors

For a collection that is not in the data set:
```{r}
simulate_random_samples(RU_starts = p$RU_starts, RU_vec = p$RU_vec, size = 100,
                          alpha_repunit = sim_spec_examples$repunit_cnt,
                          alpha_collection = sim_spec_examples$unknown_collection_error)
```

Or a repunit that is not in the data set:
```{r}
simulate_random_samples(RU_starts = p$RU_starts, RU_vec = p$RU_vec, size = 100,
                          alpha_repunit = sim_spec_examples$unknown_repunit_error,
                          alpha_collection = sim_spec_examples$coll_sub_ppn)
```
